---
title: "R Notebook"
output: html_notebook
---

---

title: "Mood Prediction with Kernel Ridge Regression and Conformal Inference"
author: "Your Name"
date: "`r Sys.Date()`"
output:
html\_document:
toc: true
toc\_depth: 3
theme: united
-------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5
)
library(caret)
library(dplyr)
library(ggplot2)
```

# Introduction

This analysis builds a **Kernel Ridge Regression** (KRR) model in an RKHS with an RBF kernel to predict individuals' mood (`Happiness`) based on questionnaire responses. We then wrap the predictions in **distribution-free conformal intervals** for valid uncertainty quantification, evaluate on a held-out test set, and compute **marginal effects** to understand feature sensitivities.

# 1. Data Loading & Preprocessing

```{r data-loading}
# Load data
data_path <- "/Users/roberto/Desktop/UNI_DATASCIENCE/Stat_learning/final_project/risposte_umore.csv"
df <- read.csv(data_path, stringsAsFactors = FALSE)
# Rename columns
colnames(df) <- c("Timestamp","Energy","SleepQuality","Hunger",
                  "PhysicalActivity","StatsLiking","Informed",
                  "BogosBinted","Anger","Fear","Happiness",
                  "Sadness","Surprise","Disgust")

# Simplify PhysicalActivity into factor
df <- df %>% mutate(
  PhysAct = case_when(
    grepl("No", PhysicalActivity, ignore.case=TRUE) ~ "No",
    grepl("leggera", PhysicalActivity, ignore.case=TRUE) ~ "Low",
    grepl("intensa", PhysicalActivity, ignore.case=TRUE) ~ "High",
    TRUE ~ "No"
  ),
  PhysAct = factor(PhysAct, levels=c("No","Low","High"))
) %>% select(-PhysicalActivity)
```

# 2. Train/Calibration/Test Split

```{r split-data}
set.seed(2025)
n <- nrow(df)
i_tr <- sample(n, size = 0.6*n)
rest <- setdiff(seq_len(n), i_tr)
i_ca <- sample(rest, size = 0.5*length(rest))
i_te <- setdiff(rest, i_ca)

features <- df %>% select(Energy, SleepQuality, Hunger,
                          StatsLiking, Informed, BogosBinted, PhysAct)
y_all <- df$Happiness

X_tr_raw <- features[i_tr,]; y_tr <- y_all[i_tr]
X_ca_raw <- features[i_ca,]; y_ca <- y_all[i_ca]
X_te_raw <- features[i_te,]; y_te <- y_all[i_te]
```

# 3. Encoding & Scaling

```{r preprocess}
dumm <- caret::dummyVars(~ ., data = X_tr_raw)
X_tr_n <- predict(dumm, X_tr_raw)
X_ca_n <- predict(dumm, X_ca_raw)
X_te_n <- predict(dumm, X_te_raw)

pp <- preProcess(X_tr_n, method = c("center","scale"))
X_tr <- predict(pp, X_tr_n)
X_ca <- predict(pp, X_ca_n)
X_te <- predict(pp, X_te_n)

# Convert to matrices
X_tr_mat <- as.matrix(X_tr)
X_ca_mat <- as.matrix(X_ca)
X_te_mat <- as.matrix(X_te)
```

# 4. Kernel Ridge Regression (KRR)

We implement KRR with an RBF kernel, selecting hyperparameters by grid-search on the calibration set.

```{r krr-grid}
# Define RBF kernel function
ingama <- function(X, Y, sigma) {
  X2 <- rowSums(X^2)
  Y2 <- rowSums(Y^2)
  D2 <- outer(X2, Y2, "+") - 2 * (X %*% t(Y))
  exp(-D2 / (2 * sigma^2))
}
# Heuristic sigma: median distance
dists <- as.vector(dist(X_tr_mat))
sigma0 <- median(dists)

sigmas <- sigma0 * c(0.5,1,2)
lambdas <- 10^seq(-3,1,length=5)
best <- list(mse=Inf)
for(s in sigmas){
  K_tr <- ingama(X_tr_mat, X_tr_mat, s)
  for(lambda in lambdas){
    A <- K_tr + lambda * diag(nrow(K_tr))
    alpha_hat <- solve(A, y_tr)
    pred_ca <- ingama(X_ca_mat, X_tr_mat, s) %*% alpha_hat
    mse_ca <- mean((pred_ca - y_ca)^2)
    if(mse_ca < best$mse) best <- list(s=s, l=lambda, a=alpha_hat)
  }
}
sigma_hat <- best$s; lambda_hat <- best$l; alpha_hat <- best$a
```

# 5. Conformal Prediction Intervals

```{r conformal}
alpha <- 0.1
pred_ca <- ingama(X_ca_mat, X_tr_mat, sigma_hat) %*% alpha_hat
nc <- abs(y_ca - pred_ca)
qlev <- (1-alpha)*(1+1/length(nc))
q_hat <- quantile(nc, probs=qlev, type=1)

predict_krr <- function(Xnew){
  mu <- ingama(as.matrix(Xnew), X_tr_mat, sigma_hat) %*% alpha_hat
  df <- data.frame(pred = as.numeric(mu))
  df$lo <- df$pred - q_hat
  df$hi <- df$pred + q_hat
  df
}
```

# 6. Evaluation on Test Set

```{r evaluate-test}
res_test <- predict_krr(X_te)
mse_test <- mean((res_test$pred - y_te)^2)
cov_test <- mean(y_te >= res_test$lo & y_te <= res_test$hi)
width_test <- mean(res_test$hi - res_test$lo)

cat(sprintf("Test MSE = %.3f\n", mse_test))
cat(sprintf("Coverage (90%%) = %.1f%%\n", 100*cov_test))
cat(sprintf("Average Width = %.3f\n", width_test))
```

# 7. Results Visualization

```{r plot-results, fig.width=7, fig.height=5}
ggplot(cbind(res_test, True=y_te), aes(x=pred,y=True)) +
  geom_point(alpha=0.6) +
  geom_errorbar(aes(ymin=lo,ymax=hi), alpha=0.3) +
  labs(title="KRR + 90% Conformal Intervals",
       subtitle=sprintf("MSE=%.3f | Cov=%.1f%% | Width=%.2f",
                        mse_test,100*cov_test,width_test),
       x="Predicted", y="True") +
  theme_minimal()
```

# 8. BogosBinted Shuffle Test

```{r shuffle-test}
shuf <- features; shuf$BogosBinted <- sample(shuf$BogosBinted)
Xsh_tr <- predict(pp, predict(dumm, shuf[i_tr,]))
Xsh_te <- predict(pp, predict(dumm, shuf[i_te,]))
A_sh   <- ingama(Xsh_tr, Xsh_tr, sigma_hat) + lambda_hat*diag(nrow(Xsh_tr))
alpha_sh <- solve(A_sh, y_tr)
pred_sh <- ingama(Xsh_te, Xsh_tr, sigma_hat) %*% alpha_sh
mse_shuf <- mean((pred_sh - y_te)^2)
cat(sprintf("Shuffle MSE = %.3f\n", mse_shuf))
```

# 9. Marginal Effects

```{r marginal-effects}
step <- 1e-3; n_tr <- nrow(X_tr_mat); p <- ncol(X_tr_mat)
meffs <- matrix(0,n_tr,p)
K_tr <- ingama(X_tr_mat, X_tr_mat, sigma_hat)
for(j in 1:n_tr){
  diff <- sweep(X_tr_mat,2,X_tr_mat[j,],"-")
  meffs[j,] <- colSums(alpha_hat * K_tr[,j] * (-diff)) / sigma_hat^2
}
avg_me <- colMeans(meffs)
quart_me <- apply(meffs,2,quantile,c(.25,.5,.75))
print(round(avg_me,8))
print(round(quart_me,8))
```

# Conclusion

This RMarkdown document implemented a custom **Kernel Ridge Regression** model with
an RBF kernel, built conformal intervals for robust uncertainty quantification,
and derived marginal effects to interpret feature contributions.

Further improvements could include:

* Adding cross-validated selection of the conformal miscoverage level  α.
* Extending to **multi-dimensional** intervals for joint prediction.
* Incorporating **time-stamped** features if longitudinal data available.

